---
title: "QM3_group_7"
author: "Keene Choy"
date: "2022-11-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## *QM3 Group 7*

# Investigating the correlation between a range of labour market factors and overall crime rate in London boroughs

## Abstract

## Data Analysis

### Installing and Loading Packages

```{r}
install.packages('readr')
library(readr)
install.packages('dplyr')
library(dplyr)
install.packages('tidyverse')
library(tidyverse)
install.packages('equatiomatic')
library(equatiomatic)
install.packages('car')
library(car)
install.packages('lmtest')
library(lmtest)
```

### Importing Cleaned Data

```{r}
crime <- read_csv('./crime.csv', show_col_types = FALSE)
head(crime)
iwp <- read_csv('./in_work_poverty.csv', show_col_types = FALSE)
head(iwp)
income<- read_csv('./income.csv', show_col_types = FALSE)
head(income)
neets<- read_csv('./neets.csv', show_col_types = FALSE)
head(neets)
emp <- read_csv('./unemployment.csv', show_col_types = FALSE)
head(emp)
```

Creating a joined dataframe containing all variables.

```{r}
reg <- left_join(crime, iwp, by = 'id')
reg <- left_join(reg, income, by = 'id')
reg <- left_join(reg, neets, by = 'id')
reg <- left_join(reg, emp, by = 'id')
head(reg)
```

### Descriptive Statistics

#### Crime Rates

```{r}
summary(reg$crime)
sd(reg$crime)
par(mfrow=c(1, 2)) 
boxplot(reg$crime, main='Crime rates in each borough each year, 2009-2016', ylab='Rate per 1000 population')
hist(reg$crime, main="Distribution of Crime Rates", xlab="Rate per 1000 population", prob=TRUE)
lines(density(reg$crime), col="red")
x <- seq(min(reg$crime), max(reg$crime), length.out=100)
y <- with(reg, dnorm(x, mean(crime), sd(crime)))
lines(x, y, col = "blue")

```

![](data:image/png;base64,iVBORw0KGgoAAAANSUhEUgAAACgAAAAaCAYAAADFTB7LAAAAcElEQVR4Xu3OwQmAQAxE0bClWYCW5N06tM6V2YPg5CjoF/JhLoHAi6iqn9eOefUbqrYvHY0cQDLyAlKRNyARmYA0ZMLRkAlGQyaU72tkAtlim7r/vJqDUDjlKBROOQyFU2icQuMUGqfQuBEaV1XPOwEx96nYACK8+wAAAABJRU5ErkJggg== "Run Current Chunk")

#### In Work Poverty

```{r}
z<- na.omit(reg$in_work_poverty)
summary(z)
sd(z)
par(mfrow=c(1, 2)) 
boxplot(reg$in_work_poverty, main='Proportion of <LLW jobs in each borough each year, 2009-2014', ylab='Percent')
hist(z, main="Distribution of Proportion of <LLW jobs", xlab="Percent", prob=TRUE)
lines(density(z), col="red")
x <- seq(min(z), max(z), length.out=100)
y <- with(reg, dnorm(x, mean(z), sd(z)))
lines(x, y, col = "blue")
```

#### Income

```{r}
summary(reg$income)
sd(reg$income)
par(mfrow=c(1, 2)) 
boxplot(reg$income, main='Resident-based median weekly earnings in each borough each year, 2009-2016', ylab='GBP')
hist(reg$income, main="Distribution of median weekly earnings", xlab="GBP", prob=TRUE)
lines(density(reg$income), col="red")
x <- seq(min(reg$income), max(reg$income), length.out=100)
y <- with(reg, dnorm(x, mean(income), sd(income)))
lines(x, y, col = "blue")
```

#### NEETs

```{r}
z<- na.omit(reg$neets)
summary(z)
sd(z)
par(mfrow=c(1, 2)) 
boxplot(reg$neets, main='Proportion of NEETs in each borough each year, 2009-2015', ylab='Precent')
hist(z, main="Distribution of Proportion of NEETs", xlab="Percent", prob=TRUE)
lines(density(z), col="red")
x <- seq(min(z), max(z), length.out=100)
y <- with(reg, dnorm(x, mean(z), sd(z)))
lines(x, y, col = "blue")
```

#### Employment

```{r}
summary(reg$employment)
sd(reg$employment)
par(mfrow=c(1, 2)) 
boxplot(reg$employment, main='Employment rates in each borough each year, 2009-2016', ylab='Percent')
hist(reg$employment, main="Distribution of employment rates", xlab="Percent", prob=TRUE)
lines(density(reg$employment), col="red")
x <- seq(min(reg$employment), max(reg$employment), length.out=100)
y <- with(reg, dnorm(x, mean(employment), sd(employment)))
lines(x, y, col = "blue")
```

### Bivariate OLS Regressions

#### Crime\~In-work Poverty

Visualizing the relationship between crime rates and proportions of \<LLW jobs

```{r}
plot(reg$in_work_poverty, reg$crime, col="blue")
abline(lm(reg$crime~reg$in_work_poverty), col="red", lwd=2)
```

Creating an OLS model and summarizing results

```{r}
modiwp<-lm(crime~in_work_poverty, data=reg)
summary(modiwp)
extract_eq(modiwp, use_coefs= TRUE)
```

Model diagnostics

```{r}
par(mfrow=c(1,2))
plot(modiwp)
plot(modiwp, 4)
residualPlots(modiwp)
shapiro.test(modiwp$residuals)
ncvTest(modiwp)
spreadLevelPlot(modiwp)
outlierTest(modiwp)
```

Linearity assumption is met (insignificant Tukey test)

Normality assumption is met (insignificant S-W normality test)

Homoscedasticity assumption is not met (significant NCV score test)

No outliers beyond Cook's Distance

Fixing heteroscedasticity using Weighted Less Square Regression:

```{r}
wt <- 1 / lm(abs(modiwp$residuals) ~ modiwp$fitted.values)$fitted.values^2
reg_wlsiwp<-na.omit(reg)
wls_modiwp <- lm(crime ~ in_work_poverty, data = reg_wlsiwp, weights=wt)
summary(wls_modiwp)
```

Re-diagnosing:

```{r}
par(mfrow=c(1,2))
plot(wls_modiwp)
plot(wls_modiwp, 4)
shapiro.test(wls_modiwp$residuals)
ncvTest(wls_modiwp)
spreadLevelPlot(wls_modiwp)
outlierTest(wls_modiwp)
```

#### Crime\~Income

Visualizing the relationship between crime rates and residence-based median weekly earnings

```{r}
plot(reg$income, reg$crime, col="blue")
abline(lm(reg$crime~reg$income), col="red", lwd=2)

plot(log(reg$income), log(reg$crime), col="blue")
abline(lm(log(reg$crime)~log(reg$income), col="red", lwd=2)
```

Creating an OLS model and summarizing results

```{r}
modinc<-lm(crime~income, data=reg)
summary(modinc)
extract_eq(modinc, use_coefs= TRUE)
```

Model diagnostics

```{r}
par(mfrow=c(1,2))
plot(modinc)
residualPlots(modinc)
plot(modinc, 4)
shapiro.test(modinc$residuals)
ncvTest(modinc)
spreadLevelPlot(modinc)
outlierTest(modinc)
```

Fixing heteroskedasticity by using Weighted Least Squares regression,

```{r}
wt <- 1 / lm(abs(modinc$residuals) ~ modinc$fitted.values)$fitted.values^2
wls_modinc <- lm(crime ~ income, data = reg, weights=wt)
summary(wls_modinc)
```

Diagnose the weighted model

```{r}
par(mfrow=c(1,2))
plot(wls_modinc)
plot(wls_modinc, 4)
residualPlots(wls_modinc)
shapiro.test(wls_modinc$residuals)
ncvTest(wls_modinc)
spreadLevelPlot(wls_modinc)
outlierTest(wls_modinc)
bptest(wls_modinc)
```

#### Crime\~NEETs

Visualizing the relationship between crime rates and proportion of NEETs

```{r}
plot(reg$neets, reg$crime, col="blue")
abline(lm(reg$crime~reg$neets), col="red", lwd=2)
```

Creating an OLS model and summarizing results

```{r}
modnee<-lm(crime~neets, data=reg)
summary(modnee)
extract_eq(modnee, use_coefs= TRUE)
```

Model diagnostics

```{r}
par(mfrow=c(1,2))
plot(modnee)
plot(modnee, 4)
residualPlots(wls_modinc)
shapiro.test(modnee$residuals)
ncvTest(modnee)
spreadLevelPlot(modnee)
outlierTest(modnee)
bptest(modnee)
```

Linearity assumption is not met (significant Tukey test)

Normality assumption is not met (significant S-W normality test)

Homoscedasticity assumption is met (insignificant NCV score test)

No outliers beyond Cook's Distance

#### Crime\~Employment

Visualizing the relationship between crime rates and employment rates

```{r}
plot(reg$employment, reg$crime, col="blue")
abline(lm(reg$crime~reg$employment), col="red", lwd=2)
```

Creating an OLS model and summarizing results

```{r}
modemp<-lm(crime ~ employment, data=reg)
summary(modemp)
extract_eq(modemp, use_coefs= TRUE)
```

Model diagnostics

```{r}
par(mfrow=c(1,2))
plot(modemp)
plot(modemp, 4)
residualPlots(modemp)
shapiro.test(modemp$residuals)
ncvTest(modemp)
spreadLevelPlot(modemp)
outlierTest(modemp)
```

Linearity assumption is met (insignificant Tukey test)

Normality assumption is not met (significant S-W normality test)

Homoscedasticity assumption is met (insignificant NCV score test)

No outliers beyond Cook's Distance
